\documentclass[12pt]{article}
\usepackage{amsfonts, epsfig}
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lfoot{\texttt{comsm0021.github.io}}
\lhead{Neural Information Processing - 1\_information\_theory\_I - Conor}
\rhead{\thepage}
\cfoot{}
\begin{document}

\section*{Worksheet} 

Many of these problems are taken from the excellent text book Cover
and Thomas. Although the questions do vary a bit in difficulty each is
worth two marks. 

\subsection*{Q1 - marginal and conditional distributions}

Work out the marginal distributions and the $x=a$ conditional distribution for
\begin{center}
\begin{tabular}{c|cc}
&$a$&$b$\\
\hline
1&$\frac{1}{3}$&$\frac{1}{6}$\\
2&0&$\frac{1}{4}$\\
3&$\frac{1}{8}$&$\frac{1}{8}$
\end{tabular}
\end{center}

\subsection*{Q2 - working out entropy}

A fair coin is flipped until the first head occurs. Let $X$ denote the
number of flips required.
\begin{enumerate}
\item Find the entropy $H(X)$ in bits. The following expressions may be useful:\begin{eqnarray}
\sum_{n=0}^{\infty} r^n&=&\frac{1}{1-r}\cr
\sum_{n=0}^{\infty}nr^n&=&\frac{r}{(1-r)^2}
\end{eqnarray}
\item A random variable $X$ is drawn according to this distribution. Find an efficient sequence of yes-no questions of the form, \lq Is $X$ contained in the set $S$?\rq{}. Compare $H(X)$ to the expected number of questions required to determine $X$.
\end{enumerate}

\subsection*{Q3 - A puzzle which lends itself to information type reasoning}

Suppose that one has $n$ coins, among which there may or may not be one counterfeit coin. If there is a counterfeit coin it will weight either less or more than the other coins. The coins are weighed using a balance.
\begin{enumerate}
\item Find an upper bound on the number of coins $n$ so that $k$ weighings will find the counterfeit coin, if any, and correctly declare it to be heavier or lighter.
\item What is the coin-weighing strategy for $k=3$ weighings and 12 coins,
\end{enumerate}

\subsection*{Q4 - Working out entropy and information}

Let $p(x,y)$ be given by $p(0,0)=p(0,1)=p(1,1)=1/3$ and
$p(1,0)=0$. Find $H(X)$, $H(Y)$, $H(X|Y)$, $H(Y|X)$, $H(X,Y)$,
$H(Y)-H(Y|X)$ and $I(X;Y)$.

\subsection*{Q5 - A question about information in the brain}

On the way.

\end{document}

